---
title: "Reproducible Research - Peer Assessment 1"
output: html_document
---
## Loading and preprocessing the data

Download and unzip the dataset with the provided link
```{r}
url <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"


download.file(url, destfile = "data.zip")
unzip(zipfile = "data.zip", exdir=".")
data <- read.csv("activity.csv")
data$date<-as.Date(data$date,format="%Y-%m-%d")
file.remove("data.zip")
file.remove("activity.csv")

```
##What is mean total number of steps taken per day?

1-Calculate daily total steps:
```{r message=FALSE}
library(dplyr)
daily_total_steps <- data %>%
  group_by(date) %>%
  summarize(daily_total_steps = sum(steps, na.rm=TRUE))
```
2-Plot a histogram
```{r}
hist(daily_total_steps$daily_total_steps, 
     main="Daily Steps Count", 
     xlab="steps", 
     col="blue",
     breaks = 10
)

```

3-Calculate mean and median steps
```{r}
mean(daily_total_steps$daily_total_steps, na.rm=TRUE)
```
```{r}
median(daily_total_steps$daily_total_steps, na.rm=TRUE)
```
What is the average daily activity pattern?
1-Time series plot
```{r}
mean_steps_perinterval <- data %>%
  group_by(interval) %>%
 summarize(mean_steps_perinterval = mean(steps, na.rm=TRUE))
library(ggplot2)
ggplot(mean_steps_perinterval, aes(x = interval, y = mean_steps_perinterval))+ geom_line()+ xlab("Interval")+ ylab("Mean number of steps ")
```
2-Maximum average number of steps and corresponding interval
```{r}
indx_max<-which.max(mean_steps_perinterval$mean_steps_perinterval)
mean_steps_perinterval[indx_max,]
```
##Imputing missing values
```{r 1-count missing values, echo=TRUE}
sum(is.na(data))
```
3-Dataset with the missing data filled in with the average number of steps
```{r}

datafill<-data
for(i in 1:ncol(datafill)){
  datafill[is.na(datafill[,i]), i] <- mean(datafill[,i], na.rm = TRUE)
}
```
4-Plot histogram
```{r}
imputed_total_steps <- datafill %>%
  group_by(date) %>%
  summarize(total_steps = sum(steps, na.rm=TRUE))
hist(imputed_total_steps$total_steps, 
     main="Daily Steps Count", 
     xlab="steps", 
     col="blue",
     breaks = 10
)
```
Mean and median are equal and higher then the values for the orginal dataset with missing values
```{r}
mean(imputed_total_steps$total_steps, na.rm=TRUE)
```
```{r}
median(imputed_total_steps$total_steps, na.rm=TRUE)
```
##Are there differences in activity patterns between weekdays and weekends?

1-Add a weekday-weekend column
```{r message=FALSE}
library(lubridate)
weekday<- c(1:5)
datafill$weekend<-ifelse((wday(datafill$date) %in% weekday), "Weekdays", "Weekend")
```
2-Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r message=FALSE}
mean_steps_week <- datafill %>%
  group_by(weekend,interval) %>%
 summarize(mean_steps = mean(steps, na.rm=TRUE))
library(ggplot2)
ggplot(data=mean_steps_week, aes(x = interval, y = mean_steps,colour=weekend)) + geom_line(aes(colour=weekend))+ xlab("Interval")+ ylab("Number of steps ")+facet_wrap(~weekend)+theme(strip.background = element_rect(fill="lightblue"))
```
